### 2014-04-01 ApTranscriptome_TR.Rmd
Continuing work to rewrite analysis working on `lm` results saved to list.

```{r responsive_transcripts}
# transcripts that differ in expression by colony only
keepthis <- function(lmitem, term = NA) {
  coefs <- coefficients(lmitem)
  coef.names <- names(coefs)
  keep <- ifelse(term %in% coef.names, TRUE, FALSE)
}

# identify sets of transcripts that include each term in final model
colony.transcripts <- RxNlmAIC[which(Map(keepthis, RxNlmAIC, term = "colonyAr") == TRUE)]
temp.lin.transcripts <- RxNlmAIC[which(Map(keepthis, RxNlmAIC, term = "val") == TRUE)]
temp.quad.transcripts <- RxNlmAIC[which(Map(keepthis, RxNlmAIC, term = "I(val^2)") == TRUE)]
col.by.temp.transcripts <- RxNlmAIC[which(Map(keepthis, RxNlmAIC, term = "colonyAr:val") == TRUE)]
col.by.temp.quad.transcripts <- RxNlmAIC[which(Map(keepthis, RxNlmAIC, term = "colonyAr:I(val^2)") == TRUE)]

# union of thermally-responsive transcripts
responsive.transcripts <- union(temp.lin.transcripts, temp.quad.transcripts)
responsive.transcripts <- union(responsive.transcripts, col.by.temp.transcripts)
responsive.transcripts <- union(responsive.transcripts, col.by.temp.quad.transcripts)
length(responsive.transcripts)

# transcripts that have a colony effect only 
colony.transcripts <- setdiff(RxNlmAIC, responsive.transcripts)

# transcripts that have colony x temperature interaction
interaction.transcripts <- union(col.by.temp.transcripts, col.by.temp.quad.transcripts)

# split responsive_transcripts into those that have same temperature response in both colonies and 
# main effect of temperature only
temperature.transcripts <- setdiff(responsive.transcripts, interaction.transcripts)
```

@@

Table 3 shows that `r round((nrow(temperature.transcripts) + nrow(interaction.transcripts))/nrow(signif.transcripts), 2)*100`% of the significant transcripts are thermally-responsive, with `r round(nrow(interaction.transcripts)/(nrow(temperature.transcripts) + nrow(interaction.transcripts)), 2)*100`% of these transcripts having a thermal-response that differs by colony.

@@

The previous section simply identified the transcripts with thermally-responsive expression. In this section, I determine the shape of the expression response to temperature for each transcript. Categories of expression response are:

```{r expression_shape, eval=TRUE, echo=TRUE}
# merge RxN results with expression values
temperature.transcripts.TPM <- TPM.dt.sub[temperature.transcripts]

>>>>

```{r responsive_transcripts}
grepFunc <- function(lmitem, term = NA) {
  coefs <- coefficients(lmitem)
  coef.names <- names(coefs)
  keep <- if(length(grep(term, coef.names)) > 0) TRUE else FALSE
}

interaction.transcripts <- RxNlmAIC[which(Map(grepFunc, RxNlmAIC, term = "colonyAr:") == TRUE)]
other.transcripts <- RxNlmAIC[setdiff(names(RxNlmAIC), names(interaction.transcripts))]
temperature.transcripts <- other.transcripts[which(Map(grepFunc, other.transcripts, term = "val") == TRUE)]
colony.transcripts <- other.transcripts[setdiff(names(other.transcripts), names(temperature.transcripts))]
rm(other.transcripts)
```

@@

The previous section identified the transcripts with thermally-responsive expression. In this section, I determine the shape of the expression response to temperature for each transcript. Categories of expression response are:

```{r expression_shape, eval=TRUE, echo=TRUE}

test.transcripts <- temperature.transcripts[1:10]

response.type <- ldply(test.transcripts, .progress = "text", RxNtype)

RxNtype <- function(lmitem) {
  # predict 
  pred.vals <- seq(from = 0, to = 38.5, by = 0.5)
  newdf <- data.frame(val = pred.vals, colony = as.factor(rep(unlist(lmitem$xlevels), each = length(pred.vals))))
  pout <- predict(lmout, newdata=newdf)
  pout <- cbind(newdf, pTPM = pout)

  max.val
  # for each colony
  for(c in levels(pout$colony)) {

    # get max val
    max.val <- pout[which(pout$pTPM == max(pout$pTPM)), "val"]
    # get min val
    min.val <- pout[which(pout$pTPM == min(pout$pTPM)), "val"]
    # determine expression shape
    exp_type <- NA
    # expression type defined as "Bimodal" if max expression below 10C and above 30C is greater than 2 standard deviations above expression at the optimum temperature (19.5C)
    if(max(pout[pout$val < 10, "pTPM"]) > (pout[pout$val == 19.5, "pTPM"] + 2*sd(pout$pTPM)) & max(pout[pout$val > 10, "pTPM"]) > (pout[pout$val == 19.5, "pTPM"] + 2*sd(pout$pTPM)))   exp_type <- "Bimodal" else {
      # expression type "High" if max expression at temperature above 30C
      if(max.val > 30) exp_type <- "High" else {
        # expression type "Low" if max expression at temperature below 10C
        if(max.val < 10) exp_type <- "Low" else
          exp_type <- "Intermediate"
        } # end else
      } # end else
    }
  }

# plot
#qplot(val, pTPM, facets = .~colony, data = pout, geom="smooth")

# merge RxN results with expression values
temperature.transcripts.TPM <- TPM.dt.sub[names(temperature.transcripts)]

@@
